Course 2 Wk 3 - Loop commands

lapply - loop over a list and evaluate a function on element
        - split is often used in conjuction with lapply

```{r}
lapply
## you will take this list of things and you want to apply a function to each element of the list and the function will return something for every single value in a new list. Always return a list

x <- list(a = 1.5, b= rnorm(10))
lapply(x, mean)
##All this is doing is taking the mean for each value of a and b
## This returned a list of length 2

x <- 1:4
lapply(x, runif) ## runif generates uniform random variables using random number generator.
        # 1st element will be 1 uniform random variable all the way down to four

lapply(x, runif, min = 0, max = 10)
# this returns random numbers only between 0 and 10


# Anonymous functions don't have a name
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x ## goal is to extract first column of each matrix
lapply(x, function(elt) elt[,1]) # the function doesn't exist except with the lapply function here. 
```




sapply: tries to simplify the result of lapply if possible
        It will return a vector instead of a list like lapply
        
- If the result is a list where every element is length 1, then a vector is returned
- If the result is a list where every element is a vector of the same length (> 1), a matrix is returned
- If it can't figure things out, a list is returned

```{r}
sapply(x, mean)
returns a vector of length two

```


apply function

- used to evaluate a function over the margins of an array (matrix is an example of an array)
        - most often used to apply a function to the rows or columns of a         matrix
        - it can be used with general arrays
        - it is not really fatser than writing a loop
        - It invovles less typing
```{r}
str(apply)
# x is an array
# margin is an integer vector indicating which margins should be retained
# FUN is a function to be applied
# ... is for other arguments to be passed to FUN

x <- matrix(rnorm(200, 20, 10))
apply(x, 2, mean)
# returns vector of length ten that has the mean of each column
# dimension 1 has 20 rows and dimension 2 has 10 columns
        # "2" eliminates the first dimension but keeps the second dimension (the 10 columns) so this preserves the columns and collapses the rows and returns a vector of length 10. You could do it so that it collapses the coloumns and preserves the rows. 

## Col/row sums and means shortcuts ##
        # rowSums = apply(x, 1, sum)
        # rowMeans = apply(x, 1, mean)
        # colSums = apply(x, 2, sum)
        # colMeans = apply(x, 2, mean)

# Calculate the 25 and 75 percentile of the matrix
apply(x, 1, quantile, probs = c(0.25, 0.75))
# this function preserves the rows
# quantile has no default function so you have to pass it the probs
# this function calculates the 25 and 75th percentile of each row

a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
apply(a, c(1,2), mean)
        # take average of the two x two matrices
        # preserves 1st and 2nd dimensions but collapses the third.
rowMeans(a, dim = 2)
## dim = 2 because this is a 2 x 2 matrix

```


mapply - multivariate funtion - applies a function in parallel over a set of arguments

Takes a multiple list and applies function over all lists

```{r}
str(mapply)
# FUN is a function to apply - number of arguments has to contain as the least the number of lists you are passing to - with 3 lists, pass three objects which means 3 arguments
# ... contains arguments to apply over
# MoreArgs
# SIMPLIFY

list(rep(1, 4), rep(2, 3), rep(3,2), rep(4,1))

# instead we can do:

mapply(rep, 1:4, 4:1)

# used to apply a function to multiple lists 

noise <- function(n ,mean, sd){
        rnorm(n, mean, sd)
}
noise(5, 1, 2)

# instead:

mapply(noise, 1:5, 1:5, 2)
## vectorizes a function

```


tapply:
used to apply a function over subsets of a vector - summarizes pieces of a vector

```{r}
str(tapply)
# X: a vector
# INDEX: is a factor or a list of factors
# FUN: function to be applied
# ... contains other arguments to be passed
# simplify: should we simplify the results?

x <- c(rnorm(10), runif(10), rnorm(10, 1)) # simulated vector of normal (rnorm) and uniform (runif) random variables. "rnorm(10, 1)" is 10 normal random variables all with a mean of 1
f <- gl(3, 10) ## gl creates a factor variable - 10 1's, 10 2's and 10 3's
f


tapply(x, f, mean) ## this allows you to take the mean of each group (in this example there are three)

tapply(x, f, mean, simplify = FALSE)

tapply(x, f, range)

```

split:
- takes a vector or other objects and splits it into groups determined by a factor or list of factors

```{r}
str(split)
# x is a vector (or list) or data frame
# f is a factor or a list of factors
# drop indicates where empty factors levels should be dropped

# if f has three levels, split will split x into three groups and then you can use sapply or tapply to pass the function to these newly created groups

x <- c(rnorm(10), runif(10), rnorm(10, 1)) # simulated vector of normal (rnorm) and uniform (runif) random variables. "rnorm(10, 1)" is 10 normal random variables all with a mean of 1
f <- gl(3, 10) ## gl creates a factor variable - 10 1's, 10 2's and 10 3's
split(x, f)

# always returns a list back

lapply(split(x, f), mean)
## tapply does the same exact thing as this



library(datasets)
head(airquality)

## calculate mean within each month
s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
## takes column means for each monthly data frame
## lapply returns a list back of vector length 3
## Ozone values NA bc there are missing values

sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
## returns a matrix instead of a list

sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
## na.rm = TRUE skips over the NA values

?gl

x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)
interaction(f1, f2)
## total of 10 levels

str(split(x, list(f1, f2), drop = TRUE))

## drop will drop the empty levels then you will be returned a list that only have observations in them
```


```{r}
library(datasets)
data(iris)
?iris

mean(iris$Sepal.Length[virginica])
head(iris)
sapply(iris$Species, unique)

lapply(iris$Sepal.length, iris$species[virginica], mean)
round(mean(iris[which(iris$Species == "virginica"),]$Sepal.Length)) 
## the mean of 'Sepal.Length' for the species virginica



apply(iris[,1:4], 2, mean)
##  returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', 'Petal.Length', and 'Petal.Width'


library(datasets)
data(mtcars)
sapply(split(mtcars$mpg, mtcars$cyl), mean)
## calculates the average miles per gallon (mpg) by number of cylinders in the car.

new <- tapply(mtcars$hp, mtcars$cyl, mean)
round(abs(new[3]-new[1]))
new
## calculates the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars

debug(ls)

```


