Course 2 Week 4


str: Compactly display the internal structure of an R object - alternative to 'summary'
- produces one line of output

Answers: What's in this object?


```{r}
str(lm)

x <- rnorm(100, 2, 4)
summary(x)
str(x) ## gives you an idea of what the data looks like

f <- gl(40, 10)
summary(f)
str(f) ## much more efficient to pull data info

library(datasets)
head(airquality)
str(airquality)

s <- split(airquality, airquality$Month) ## split the airquality dataset by month
str(s) ## this is a list that contains five different data frames


```


Simulation: ****** NEED TO DO MORE RESEARCH ON THESE

functions for probability distribution:
1. rnorm - simulates random normal variables with given mean and sd
2. dnorm - evaluate the Normal probability densiy at a point (or vector of points)
3. pnorm - evaluate the cumulative function for a normal dist
4. rpois - generate random Poisson variates with a given rate
        - WHAT IS A POISSON variates
        - a discrete probability distribution that expresses the probability of a given number of events occurring in a fixed interval of time and/or space if these events occur with a known average rate and independently of the time since the last event.[1]


- d = for density
- r = for random number generation
- p = cumulative distribution
- q = for quantile function

Example:
rpois
dpois
ppois
qpois
```{r}
str(dnorm)
str(rnorm)
x <- rnorm(10)

x2 <- rnorm(10,20,2)

x2

## any time you simulate random variables, you need to set seed
set.seed(1) # without this, the variables are not random, but when you set the seed, you ensure reproducibility 

# allows you to reproduce if you use set the seed - you will want to do this because you want to be able to get same results to find/solve any errors

rpois(10,1) ## integer

ppois(2, 2) # Cumulative 
# what is the probability that the poisson distribution is less than or equal to 2 with a rate of 2

```

Generating Random numbers from a linear numbers

Suppose we want to simulate the following linear model:

y = beta(0) + beta(1)x + e
where e = N(0.2^2),
assume x ~ N(0.1^2), beta(0) = 0.5 and beta(1) = 2

```{r}
set.seed(20)
x <- rnorm(100) ## generate x, the predictor
e <- rnorm(100, 0, 2) # normal distribution 
y <- 0.5 + 2 + x + e # add them all together after multiplyinig the coefficients
summary(y)
plot(x, y) ## y simulated from linear model



## For binary:

set.seed(10)
x <- rbinom(100, 1, 0.5) ## generate x, the predictor
e <- rnorm(100, 0, 2) # normal distribution 
y <- 0.5 + 2 + x + e # add them all together after multiplyinig the coefficients
summary(y)
plot(x, y) ## x variable is binary - plot shows there is a clear linear trend when going from zero to 1


```


Simulate a Poisson model

Y ~ Poisson(mu)
log(mu) = beta(0) + beta(1)x
and beta(0) = 0.5 and beta(1) = 0.3. We need to use rpois function for this

```{r}
set.seed(1)
x <- rnorm(100)
log.mu <- 0.5 + 0.3 * x ## linear predictor
## follows a linear model
y <- rpois(100, exp(log.mu)) ## exp(log.mu) gives exponential of log mean
summary(y)
plot(x, y) # linear relationship between x and y




```


Random Sampling:

"sample" function draws random samples from that vector - can create arbitrary distribution
```{r}
set.seed(1)
sample(1:10, 4)

sample(1:10, replace = TRUE) # sampling 1-10 but replacing so you can get repeats




```
```{r}
set.seed(1)
rpois(5, 2)


library(swirl)

```

