What is R?

S is a langauage developed in 1976. In 1988 the system was rewritten in 
C and begins to resemble today stuff. Code was then rewritten again in 1998 and resembles the code we use today. R is thing of S.

Fundamentals of S language have not changed much since 1998

R was created in 1991. R can only be modified by those in the core group

Syntax is very similar to S. Semantics are superficially similar to S

Smart way to ask help:
"R 3.0.2 lm() function on Windows -- seg fault on large data frame"

Five atomic classes of objects:
1. Character
2. Numeric - numbers
3. Integer - 1L gives you an integer instead of numeric 
4. Complex
5. Logical
Inf = infinity
NaN - Not a Number, can also be thought of as a missing value

Vector must contain objects of same class
"list" is the execption and can contain objects of different classes

Attributes:
- Names, dimnames
- dimensions
- class
- length
- other user-defined attributes/metadata
can be accesssed using attributes()

Cancatinating:
x <- c(1, 2)
Using vector():
y <- vector(numeric, 1:10)

Vector Coercion:
y <- c("a", TRUE) ## character
y <- c(TRUE, 2) ## numeric
```{r}
# Course 2, Wk 1 HW #

y <- c(2, T)
y
class(y)

 x <- c(1,3, 5) 
 y <- c(3, 2, 10)
 rbind(x, y)


x <- list(2, "a", "b", TRUE)
x[[1]]


x <- c(3, 5, 1, 10, 12, 6)
x[x %in% 1:5] <- 0
x[x<6] <- 0
## these set all numbers less than 6 = 0


colnames(hw1_data)
hw1_data[c(1,2),]
hw1_data[c(152, 153),]
hw1_data[47,1]

miss <- is.na(hw1_data[, "Ozone"])  ## A vector of TRUE/FALSE
sum(miss)


mean(na.omit(hw1_data$Ozone)) ##takes mean w/o NA values



nrow(hw1_data)

new.Ozone <- hw1_data$Ozone[hw1_data$Ozone>31]
new.Ozone

new_hw1 <- subset(hw1_data, Ozone > 31 & Temp > 90)
mean(new_hw1$Solar.R)
## Takes mean when Ozone > 31 and Temp is > 90

tapply(hw1_data$Temp, hw1_data$Month, mean, na.rm = TRUE)
## takes mean Temp by each month
tapply(hw1_data$Ozone, hw1_data$Month, max, na.rm = TRUE)
### Takes maximum ozone value by each month
```

Need to be careful about what type of coercion you will get

Converting:
x <- 0:6
class(x)
"integer"
as.numeric(x)
as.logical(x) ## FALSE T T T T T T bc 0=False & 0< = True

as.complex

## List
x <- list(1, "a", TRUE, 1 + 4i)
x
[[1]]
[1] 1

[[2]]
[1] "a"

Etc...
Elements of the list will have double brackets

## Matrices

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m

m1 <- 1:10
m1
dim(m1) <- c(2, 5)
m1
```

##Cbind and Rbind:
```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
```

## Factors - represents categorical data:
unordered (male/female) or ordered (ranked such as associate, assistant, etc)

treated by lm() and glm()
Factors are self-describing
  Coding for label is built-in

```{r}
x <- factor(c("yes", "no", "yes"))
x
table(x)
unclass(x) ##brings it down to integer vector

```
The order of the levels can be set using "levels" argument to facctor()
```{r}
x <- factor(c("yes", "no", "yes"),
            levels = c("yes", "no")
```

## Missing Values

is.na used to test objects if they are NA
is.nan used to test for NaN

NA values can have class
BUT NaN is considered to be NA but NA value is not necessarily an NaN value

## Data Frames

used to store tabular data - make up a lot of what we use in statistics

data frames are represented as a special type of list
  each column has to be same length but not same type
  Have special attribute called "row.names"
  created using "read.table()" OR "read.csv"
  Can be converted to matrix using data.matrix()
      Need to be careful with coercion here

```{r}
x <- data.frame(foo = 1:4, bar = c(T, F, T, T))
x
```

## Names:

```{r}
x <- 1:3
names(x) <- c("foo", "bar", "north")
x
```

dimnames() passes names to list


## Reading Tabular Data:

read.table()
read.csv()
    These function read text files and return a data frame in R
  
readLines for reading lines of a text file
source for reading in R code files
dget for reading in R code files

read.table:
   - file
   - header: logical indicating if the file has a header line
   - sep: string indicates how columns are separated
   - colClasses: char vector indicating class of coulmn
   - nrows: number of rows in the dataset
   - comment.char: indicated comment charachter
   - skip: number of lines to skip from beginning
   - stringsAsFactors - should character variables be coded as            factors? Default is TRUE

## Textual Data Formats

dump()

dput() - use dget to read in the data
These Preseve the metadata


## Subsetting
[] - returns object of same class as the original, if you subset a list, a list will be returned. Can be used to select more than one element

[[]] - used to extract elements of a list or data frame - can only be used to extract a single element

```{r}
x <- c("a", "b", "c", "abc")
x[1]

x[x>"a"]

u <- x > "a" ##subset to see which elements are greater than "a"
u

```

## Subsetting lists

```{r}
x <- list(foo = 1:4, bar = 0.5)
x[1]
x[[1]]

x$bar

x["bar"] ## list with element bracket in it

x[c(1, 2)] ## extracting multiple elements

x <- list(a = list(10, 12, 14), b = c(3, 5))
x[[c(1, 3)]]
##extracts the third element of the first list
```


## Subsetting: Removing NA values

```{r}
x <- c(1, 2, NA, 4, NA, 5, NA, 7, 8, NA)
bad <- is.na(x)
bad
x[!bad] ## pulls out which ones aren't NA

y <- c(NA, "a", NA, "b", NA, "c", NA, NA, NA, NA)
good <- complete.cases(x, y) ## will put out only variables from both that are not NA - will match and substitute out all missing variables
good
x[good]
y[good]

```



```{r}
install.packages("swirl")
library(swirl)

install_from_swirl("R Programming")
swirl()


```
