Course 3 Wk 4 Notes of Getting and Cleaning Data

Editing text data sets
```{r}
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl, destfile = "./data/cameras.csv")
cameraData <- read.csv("./data/cameras.csv")
names(cameraData)

tolower(names(cameraData)) ## tolower function makes all letters lower case

## toupper command makes all letters upper case

#For separating values by periods:
splitNames = strsplit(names(cameraData),"\\.")
splitNames[5]

## Quick asie on lists:
myList <- list(letters = c("A", "B", "C"), numbers = 1:3, matrix(1:25, ncol = 5))
head(myList) ## creating it by passing letters, numbers and matrix

myList[1]

## Remove all the periods:
splitNames[[6]][1]

## Create variable to pull first value out of the list:
firstElement <- function(x){x[1]}
sapply(splitNames, firstElement) ## use sapply the split names (which is a list) to firstElement to remove the periods from names in data frames.

fileUrl1 = "https://dl.dropboxusercontent.com/u/7710864/data/reviews-apr29.csv"
fileUrl2 = "https://dl.dropboxusercontent.com/u/7710864/data/solutions-apr29.csv"
download.file(fileUrl1, destfile = "./data/reviews.csv")
download.file(fileUrl2, destfile = "./data/solutions.csv")
reviews = read.csv("./data/reviews.csv"); solutions <- read.csv("./data/solutions.csv")

head(reviews, 2)
names(reviews)
sub("_","",names(reviews),) ## saying that every time there is an underscore, replace it with ""

## Using gsub to replace multiple instances - sub only replaces the first instance:
testName <- "this_is_a_test"
sub("_","",testName)
gsub("_","",testName) ## replaces all instances


## Search for specific variables/value names: grep() & grepl()
grep("Alameda", cameraData$intersection)
table(grepl("Alameda", cameraData$intersection))
## returns a vector true if value appears, false, if it doesn't

## Subset to find all cases where Alameda appear in the data and if it doesn't, then I want to subset differently
cameraData2 <- cameraData[!grepl("Alameda", cameraData$intersection),]
cameraData2 ## subset data based on certain searches you want to find

grep("Alameda", cameraData$intersection,value=TRUE)
## Returns value where Alameda appears
grep("JeffStreet", cameraData$intersection)
## Look for values that don't appear - will show integer(0)
length(grep("JeffStreet", cameraData$intersection))


library(stringr)
nchar("Jeffrey Leek")

substr("Jeffery Leek", 1, 7) ## 1 refers to where you want to start pulling the data and 7 is the character you want to end at

paste("Jeffery", "Leek") ## combine strings of text together

paste0("Jeffery", "Leek")

str_trim("Jeff      ")



```
Important points about text in data sets:
- Names of variables should be:
1. All lower case when possible
2. Descriptive (Diagnosis vs Dx)
3. Not duplicated
4. Not have underscores or dots or white spaces
- Variables with character values:
1. Shoudl usually be made into factor variables
2. Should be descriptive: (Use True/False instead of 0/1 and Male/Female versus 0/1 or M/F)




Regular Expressions:
```{r}
## literals and metacharacter (define its grammar) the latter specifcy what we want much better

^i think ## matches all of the lines that start with i think, but would not match if it was in the middle or at the end of the line

^[Ii] am ## looks at lower or upper case i and the literal "am"

^[0-9][a-zA-Z] ## start at beginning of line and then searches for a letter following - it will look for a line that starts with a # and then a letter following immediately afterwards 

## "." is used to refer to any character:
9.11 ## will search for lines that have a 9 and an 11 separated by one character

fire|flood ## matches any line that includes either fire or flood

## ^[Gg]ood|[Bb]ad




```


Working with dates:

```{r}

d1 = date()
d1

d2 = Sys.Date()
d2
class(d2)


## Reformt dates:
format(d2,"%a %b %d")
## These are abbreviations for dats

weekdays(d2)

months(d2)

julian(d2)

install.packages("lubridate")
library(lubridate); ymd("20140108") ## formats it into dates

ymd_hms("2011-08-03 10:15:03")

## tz =    This function sets the time zone

## Lubridate tutorial 
## Want the classes in "POSIXct" and POSITct
```


Open gov sites:

un.org
data.gov
data.gov.uk
data.gouv.fr
data.gov/opendatasites

Gapminder has data about development and human health

Survey data from US - asdfree.com teaches you how to use surveys in R


```{r}
## Quiz Wk 4

fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileUrl, destfile = "./data/Idaho.csv")
Idaho <- read.csv("./data/Idaho.csv")

names(Idaho)
splitNames2 = strsplit(names(Idaho),"wgtp" )
splitNames2[[123]]
## another option: strsplit(names(data), "\\wgtp")[123]


## Question #2
country_GDP <- read.csv("./data/GDP", skip = 5, nrows = 190, stringsAsFactors = FALSE, header = FALSE)

names(country_GDP)
head(country_GDP, 30)
str(country_GDP)


country_GDP$V5 <- as.numeric(gsub(",","",country_GDP$V5))
country_GDP$V5
mean(country_GDP$V5, na.rm = TRUE)

## Another way
gdpData <- country_GDP[,c(1,2,4,5)]
head(gdpData)
colnames(gdpData) <- c("CountryCode", "Rank", "CountryName", "GDP")
mean(as.numeric(gsub(",", "", gdpData$GDP)), na.rm = TRUE)

## Question # 3

grep("^United",gdpData$CountryName)
length(grep("^United",gdpData$CountryName))


## Question # 4
fileEdu <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
download.file(fileEdu, destfile = "./data/Edu")
dataEdu <- read.csv("./data/Edu", stringsAsFactors = FALSE)
head(dataEdu)

mergedData <- merge(gdpData, dataEdu, all = TRUE, by = c("CountryCode"))
mergedData
length(grep("[Ff]iscal year end(.*)+June", mergedData$Special.Notes))


## Question #5

install.packages("quantmod")
library(quantmod)
amzn = getSymbols("AMZN",auto.assign=FALSE)
sampleTimes = index(amzn)
head(sampleTimes)

sampleTimes = ymd(sampleTimes) ## put it in a date format that can be read
length(grep("^2012", sampleTimes))
Y2012 <- subset(sampleTimes, year(sampleTimes) == 2012)
## Find out number of Mondays in this subset
length(which(wday(Y2012, label = T) == "Mon")) ## [1] 47


library(swirl)


```



Question 1:
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf

Apply strsplit() to split all the names of the data frame on the characters "wgtp". What is the value of the 123 element of the resulting list?


Question 2:
Load the Gross Domestic Product data for the 190 ranked countries in this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv

Remove the commas from the GDP numbers in millions of dollars and average them. What is the average?

Original data sources:

http://data.worldbank.org/data-catalog/GDP-ranking-table




Question 3:
In the data set from Question 2 what is a regular expression that would allow you to count the number of countries whose name begins with "United"? Assume that the variable with the country names in it is named countryNames. How many countries begin with United?




Question 4:
Load the Gross Domestic Product data for the 190 ranked countries in this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv

Load the educational data from this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv

Match the data based on the country shortcode. Of the countries for which the end of the fiscal year is available, how many end in June?

Original data sources:

http://data.worldbank.org/data-catalog/GDP-ranking-table

http://data.worldbank.org/data-catalog/ed-stats




Question #5
You can use the quantmod (http://www.quantmod.com/) package to get historical stock prices for publicly traded companies on the NASDAQ and NYSE. Use the following code to download data on Amazon's stock price and get the times the data was sampled.


