Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a
state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num).
The function reads the outcome-of-care-measures.csv file and returns a character vector with the name
of the hospital that has the ranking specified by the num argument. 

For example, the call:
rankhospital("MD", "heart failure", 5)
would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate
for heart failure. The num argument can take values "best", "worst", or an integer indicating the ranking
(smaller numbers are better). If the number given by num is larger than the number of hospitals in that
state, then the function should return NA. Hospitals that do not have data on a particular outcome should
be excluded from the set of hospitals when deciding the rankings.
```{r}
rankhospital <- function(state, outcome,num='best') {
        if(is.character(num)){
        if (num %in% c('best','worst')==FALSE){
            stop('Invalid Rank')
        }
    
    data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
    
    if (state %in% data$State ==FALSE) { stop("invalid state")}
    
    if (outcome %in% c("heart failure","heart attack","pneumonia")==FALSE){
        stop("invalid outcome")}
    
   
 
    
    if (state %in% data$State & outcome %in% c("heart failure","heart attack","pneumonia")){
     if(outcome=="heart attack"){
         a<-data[,c(2,7,11)]
     }
     
     if(outcome=="heart failure"){
         a<-data[,c(2,7,17)]
     }
     
     if(outcome=="pneumonia"){
         a<-data[,c(2,7,23)]
     }
     
     b<-split(a,a$State)
     c<-b[[state]]
     c<-c[,c(1,3)]
     colnames(c)<-cbind('Hospital.Name',outcome)
     c[,2]<-as.numeric(c[,2])
     
     
     if (num=='best'){
                              
     d<- c[which(c[,2]==min(c[,2],na.rm=TRUE)),]

     hospitals<-d$Hospital.Name
     hospital<-sort(hospital)[1]
     
     }
     
     if (num=='worst'){
         
         d<- c[which(c[,2]==max(c[,2],na.rm=TRUE)),]
         
         hospitals<-d$Hospital.Name
         hospital<-sort(hospitals)[1]
         
     }
     
    if(is.numeric(num) & num < length(c$Hospital.Name)){
        
        
        l<-c[!is.na(c[,2]),]
        ll<-unique(sort(l[,2]))
        
         m<-c()
        
        for ( i in 1:length(ll)){
            temp<-which(l[,2]==ll[i])
            m<-rbind(m,l[temp,])
        }
        
        k<-m[num,2]
        p1<-which(m[,2]<k)
        p<-which(m[,2]==k)
        
        num2<-num-length(p1)
        
        hospitals<-sort(m[p,1])
        
        hospital<-hospitals[num2] 
        
    }
     
    if (is.numeric(num)){
        if (num > length(c$Hospital.Name)){ 
            return(NA)}}

    }
    
    hospital   
}
    
    
}


```


```{r}
rankhospital <- function(state, outcome, num = "best") {

    ## Read the outcome data
    dat <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
    ## Check that state and outcome are valid
    if (!state %in% unique(dat[, 7])) {
        stop("invalid state")
    }
    switch(outcome, `heart attack` = {
        col = 11
    }, `heart failure` = {
        col = 17
    }, pneumonia = {
        col = 23
    }, stop("invalid outcome"))
    dat[, col] = as.numeric(dat[, col])
    df = dat[dat[, 7] == state, c(2, col)]
    df = na.omit(df)
    nhospital = nrow(df)
    switch(num, best = {
        num = 1
    }, worst = {
        num = nhospital
    })
    if (num > nhospital) {
        return(NA)
    }
    ## Return hospital name in that state with the given rank 30-day death rate

    o = order(df[, 2], df[, 1])
    df[o, ][num, 1]
}


```

